---
title: "hw1"
author: "Renjie Liu"
date: "1/12/2021"
output: html_document
---

*1.	For each of the situations below, identify the response and explanatory variables, variable types, and the generalized linear model that is well-suited to model the data. Make sure to justify your answer. Write down the linear predictor as well as the link function.
*The effect of age, sex, height, daily food intake and minutes of daily exercise on a person’s weight.
<br> Response: person’s weight type: continuous <br>
Explanatory variables: <br>
 age: discrete (recorded in years) <br>
 sex:nominal (male and female) <br>
 height: continuous (measured in feet) <br>
 daily food intake:continuous (measured in calories) <br>
 minutes of daily exercise: discrete (measured in mintues <br>
 Linear predictor:  xTiβ=β0+ β1age + β2sex+ β3heigt + β4dailyfoodintak + β5minutesofdailyexercise <br>
 Link function : g(μi)=μi=xTiβ <br>




•	The effect of sex, age, GPA, major, prior years of work experience, and prior income levels on whether a full-time graduate students finds employment upon graduation
Response: full time graduate student finds employment upon graduation or not. Type = cagegorical (yes or no).
Explanatory variables:<br> 
sex: nominal <br>
age: discrete <br>
GPA: continous (measured as number) <br>
major: nominal  (not number) <br>
Prior years of work experience: discrete (int) <br>
prior income levels: ordinal  (measured in lvls ) <br>
Linear Predictor: xTiβ=β0+β1Sex+β2Age+β3GPA+β4Major+β5WorkExp+β6IncomeLevel <br>
Link function g(μi)=ln(μi1−μi)=xTiβ <br>



•	The number of mortgage loan defaults in a given year by different counties across the United States. For each household/borrower information on income, loan interest rate, age, debt, loan to value at origination are available.
Response: number of mortgage loan defaults in a year by different counties across the us.  Type: discrete counts.
Explanatory:  <br>
borrowers’ income: continueous (measure in numbers) <br>
interest rate: contineous  (measure in numbers) <br>
age: discrete (int) <br>
debt: contineous (number) <br>
loan to value : contineous (number) <br>
Linear Predictor:  xTiβ=β0+β1Income+β2InterestRate+β3Age+β4Debt+β5LoanToValue <br>
Link function g(μi)=ln(μi)=xTiβ <br>

 
 

2.	Assume that the linear predictor, xTiβ, takes the values between −3−3 and 33 equally spaced by 0.010.01. If the link function is the natural logarithm, log(μi)=xTiβlog(μi)=xiTβ, then compute and plot the mean response, μiμi, against the linear predictor, xTiβxiTβ.

```{r}
#the mean response is the inverse of the logit function with respect to the linear predictor

x<- seq(-3,3,0.01)
y<- exp(x)
plot(x,y)

```




3.	Assume that the linear predictor, xTiβxiTβ, takes the values between −5−5 and 55 equally spaced by 0.010.01. If the link function is logit, log(μi1−μi)=xTiβloga(μi1−μi)=xiTβ, then compute and plot the mean response, μiμi, against the linear predictor, xTiβxiTβ.
```{r}
x<- seq(-5,5,0.01)
y<- exp(x)/(1+exp(x))
plot(x,y)
```



4.	Assume that the linear predictor, xTiβxiTβ, takes the values between 0.50.5 and 55 equally spaced by 0.010.01. If the link function is negative inverse, −μ−1i=xTiβ−μi−1=xiTβ, compute and plot the mean response, μiμi, against the linear predictor, xTiβxiTβ.
```{r}
x<- seq(0.5,5,0.01)
y<- 1/(-x)
plot(x,y)
```


