---
title: "HW4-Renjie Liu"
author: "Renjie Liu"
date: "2/8/2021"
output: html_document
---
HW4
1.	Estimate a generalized linear model assuming the response follows a Poisson distribution, the mean response is linked to the linear predictor by a log function. Use the dataset ‘’z1037.poisson-loglink.RData’’ to estimate the parameters in the linear predictor by maximum likelihood estimation.
•	First, estimate the parameters of the GLM with Poisson response and log link function by modifying the Newton-Raphson algorithm provided above for the regression with linear link function. Plot the values of the parameters for every iteration. Compute and show the values for W, z, XTWX, and XTWz on the last iteration.
```{r}
library(reshape2)
# ITERATIONS
load("z1037.poisson-loglink.RData")
dta <- data.frame(ite=numeric(),b0=numeric(),b1=numeric(),err=numeric(),stringsAsFactors=FALSE)

y <- matrix(regdta$y,ncol=1)
X <- matrix(c(regdta$x1,regdta$x2), nrow = length(regdta$y), ncol=2, byrow = FALSE)

b0 <- c(1,1)
tol <- 1e-5
err <- 1e10
ite <- 0

while (err>tol){
  
  var.y <- (X %*% b0)
  W <- diag(as.numeric(exp(X %*% b0)^2/var.y),nrow=length(var.y),ncol=length(var.y))
  z <- X %*% b0 + ((y-exp(X %*% b0))*1/exp(X %*% b0))
  
  b1 <- solve(t(X) %*% W %*% X) %*% (t(X) %*% W %*% z)
  
  err <- max(abs(b1-b0))
  dta <- rbind(dta,data.frame(ite,b0,b1,err,stringsAsFactors=FALSE))
  
  b0 <- b1
  ite <- ite + 1
}

```

```{r}
W
z
t(X) %*% W %*% X
t(X) %*% W %*% z
b.no <- rep(c(paste0("b_",1),paste0("b_",2)),nrow(dta)/nrow(b0))
dta <- cbind(dta,b.no)
dta <- dta[,c("ite","b.no","b0")]
dta <- dcast(dta,ite~b.no,value.var="b0")
dta
par(mfrow=c(1,2), mar=c(2,5,2,1),oma=c(0,0,3,0))
plot(dta$ite,dta$b_1,ylim=c(min(dta$b_1),max(dta$b_1)),xlab="No of Iterations",ylab="beta0",main="",pch=16,col="red",lty=2)
plot(dta$ite,dta$b_2,ylim=c(min(dta$b_2),max(dta$b_2)),xlab="No of Iterations",ylab="beta1",main="",pch=16,col="red",lty=2)

mtext("Newton-Raphson Method - Poisson (Log)",outer=TRUE,cex=1.5)
```
•	Second, estimate the parameters of the Poisson regression with log link function using the glm function in stats R package.
```{r}
model1 <- glm(y~x2,family = poisson("log"),data = regdta)
model1$coefficients
#Yes I got the same answer. 
```


2.	Estimate a generalized linear model assuming the response follows a Normal distribution, the mean response is linked to the linear predictor by an identity function. Use the dataset ‘’z1037.normal-identity.RData’’ to estimate the parameters in the linear predictor by maximum likehood estimation.
•	First, estimate the parameters of the GLM with Normal response and identity link function by modifying the Newton-Raphson algorithm provided above for the regression with linear link function. Plot the values of the parameters for every iteration. Compute and show the values for WW, zz, XTWXXTWX, and XTWzXTWz on the last iteration.
```{r}
load("z1037.normal-identity.RData")
dta <- data.frame(ite=numeric(),b0=numeric(),b1=numeric(),err=numeric(),stringsAsFactors=FALSE)

y <- matrix(regdta$y,ncol=1)
X <- matrix(c(regdta$x1,regdta$x2),nrow = length(regdta$y), ncol=2, byrow = FALSE)

b0 <- c(7,1)
tol <- 1e-5
err <- 1e10
ite <- 0

while (err>tol){

    var.y <- var(y)
    W <- diag(as.numeric(1/var.y),nrow=length(regdta$y),ncol=length(regdta$y))
    z <- y

    b1 <- solve(t(X) %*% W %*% X) %*% (t(X) %*% W %*% z)

    err <- max(abs(b1-b0))
    dta <- rbind(dta,data.frame(ite,b0,b1,err,stringsAsFactors=FALSE))
    
    b0 <- b1
    ite <- ite + 1
}
```

```{r}
W
z
t(X) %*% W %*% X
t(X) %*% W %*% z
b.no <- rep(c(paste0("b_",1),paste0("b_",2)),nrow(dta)/nrow(b0))
dta <- cbind(dta,b.no)
dta <- dta[,c("ite","b.no","b0")]
dta <- dcast(dta,ite~b.no,value.var="b0")
dta
par(mfrow=c(1,2), mar=c(2,5,2,1),oma=c(0,0,3,0))
plot(dta$ite,dta$b_1,ylim=c(min(dta$b_1),max(dta$b_1)),xlab="No of Iterations",ylab="beta0",main="",pch=16,col="red",lty=2)
plot(dta$ite,dta$b_2,ylim=c(min(dta$b_2),max(dta$b_2)),xlab="No of Iterations",ylab="beta1",main="",pch=16,col="red",lty=2)

mtext("Newton-Raphson Method - Gaussian (Linear)",outer=TRUE,cex=1.5)
```

•	Second, estimate the parameters of the model using the GLM function in the R package.
```{r}
model2 <- glm(y~x2,family = gaussian("identity"),data = regdta)
model2$coefficients
```

•	Do you get the same answer?
```{r}
#Yes I got the same answer.
```

•	State the mathematical representation of the the least squares estimator from a standard linear regression in matrix form.
 
 $\beta = (X^{T}X)^{-1}X^{T}Y$


•	Compute the estimate for the beta coefficients using the least squares estimator.
```{r}

betacoef <- solve(t(X) %*% X) %*% t(X) %*% y
betacoef

```

• Do you get the same answer?
```{r}
#Yes I got the same answer
```


